Report on a Safety Checker for Petri nets

### 2014-02-01 paper/workspace/main-example-filip.tex
Added version of paper draft with Filip's example

The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].

@@

* Property: Thread 1 and Thread 2 are never in the critical section at the same time.

@@

\begin{center}
  \includegraphics[width=12cm]{lamport}
\end{center}

\begin{verbatim}
* Property P: cs1 + cs2 <= 1
\end{verbatim}

@@

    id1        = 1 + u1 - u2
    ^            ^   ^    ^
    |            |   |    |
    |            |   |    |
    |            |   |    |
    |            |   |   # of tokens taken from nc
    |            |   |    
    |            |  # of tokens given to nc
    |            |
    |          initial number of tokens in nc
    |
    number of tokens in nc

    cs1        = 0 - u1      + u3
    idle1      = 1 + u1 - u2
    req1       = 0      + u2 - u3
    nid1       = 0 - u1 + u2
    id2        = 1                     + v2           + v5 - v6
    cs2        = 0                               + v4 - v5
    idle2      = 1                          + v3 + v5      - v6
    req2       = 0                - v1           - v4      + v6
    after_you2 = 0                + v1 - v2
    await2     = 0                     + v2 - v3
\end{verbatim}

@@

    id1        >= 0
    cs1        >= 0
    idle1      >= 0
    req1       >= 0
    nid1       >= 0
    id2        >= 0
    cs2        >= 0
    idle2      >= 0
    req2       >= 0
    after_you2 >= 0
    await2     >= 0
    u1         >= 0
    u2         >= 0
    u3         >= 0
    v1         >= 0
    v2         >= 0
    v3         >= 0
    v4         >= 0
    v5         >= 0
    v6         >= 0

@@

    cs1 + cs2  >= 2

@@

  \end{verbatim}
\begin{align*}
  C =& \left( \bigwedge_{s \in S}
    s = M_0(s) + \sum_{t \in ^\bullet s} t - \sum_{t \in s^\bullet} t
  \right) \land \\
  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

@@

  - Assignment A:
    id1        = 0
    cs1        = 1
    idle1      = 0
    req1       = 0
    nid1       = 1
    id2        = 0
    cs2        = 1
    idle2      = 0
    req2       = 0
    after_you2 = 0
    await2     = 0
    u1         = 0
    u2         = 1
    u3         = 1
    v1         = 0
    v2         = 0
    v3         = 0
    v4         = 1
    v5         = 0
    v6         = 1

@@

    - Trap implications:
                       v4              v3                 u2
                  ------------    --------------    --------------     
    id1        => (cs2 \/ id1) /\ (id1 \/ idle2) /\ (req1 \/ nid1)
    ^              ^      ^        ^      ^          ^       ^
    |              |      |        |      |          |       |
    |              |     id1 \in S |     idle2 \in S |      nid1 \in S
    |              |               |                 |
    |             cs2 \in S       id1 \in S         req1 \in S
    |
    id1 \in S

    cs1        => idle1 \/ id1
    idle1      => req1 \/ nid1
    req1       => cs1 \/ id2
    nid1       => (after_you2 \/ nid1) /\ (idle1 \/ id1)
    id2        => req2 /\ (cs1 \/ id2)
    cs2        => (idle2 \/ id2)
    idle2      => req2
    req2       => (cs2 \/ id1) /\ (cs2 \/ id1)
    after_you2 => id2 \/ await2
    await2     => id1 \/ idle2

    - At least one of the initially marked places belongs to S:
    id1 \/ id2 \/ idle1 \/ idle2

    - None of the marked places in A belongs to S:
    ~cs1 /\ ~cs2 /\ ~nid1
\end{verbatim}

@@

\begin{verbatim}
  - Assignment A_\theta:
    id1        = true
    cs1        = false
    idle1      = false
    req1       = true
    nid1       = false
    id2        = true
    cs2        = false
    idle2      = false
    req2       = true
    after_you2 = true
    await2     = false
  - Refinement constraint \delta:
    id1 + id2 + req1 + req2 + after_you2 >= 1
    ^     ^     ^      ^      ^
    |     |     |      |      |
    ---------------------------
      |
      S = {iq1, iq2, req1, req2, after_you2}, therefore \delta excludes
      assignment A in the next iteration

@@

\end{verbatim}
\begin{align*}
  C_\theta =& \left( \bigwedge_{s \in S} \left( s \Rightarrow
        \bigwedge_{t \in s^\bullet} \bigvee_{p \in t^\bullet} p
    \right) \right) \land \\
  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
\end{align*}
\begin{verbatim}

@@

\end{verbatim}
\begin{align*}
  & \delta = \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
\end{align*}

>>>>

The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
We present the method Safety by example on a modified Petri net from [Esparza1997].

@@

* Property: z is never marked

@@

\begin{center}
  \includegraphics[width=6cm]{example-filip}
\end{center}

\begin{verbatim}
* Property P: z = 0
\end{verbatim}

@@

    x = 1 - t
    ^   ^   ^
    |   |   |
    |   |   |
    |   |   |
    |   |   |
    |   |   |
    |   |  # of tokens taken from x
    |   |
    |   initial number of tokens in x
    |
    number of tokens in x

    y = 0
    z = 0 + t
    w = 0 + t
\end{verbatim}

@@

    x >= 0
    y >= 0
    z >= 0
    w >= 0
    t >= 0

@@

    z >= 1

@@

  \end{verbatim}
\begin{align*}
  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
  \right) \right) \land \\
  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

@@

  - Assignment A:
    x = 0
    y = 0
    z = 1
    w = 1
    t = 1

@@

    - Trap implications:
              t
         -----------
    x => y \/ z \/ w
    ^    ^    |    |
    |    |    |    w \in S
    |    |    |
    |    |    z \in S
    |    |
    |    y \in S
    |
    x \in S

    y => y \/ z \/ w
    z => true
    w => true

    - At least one of the initially marked places belongs to S:
    x

    - None of the marked places in A belongs to S:
    ~z /\ ~w
\end{verbatim}

@@

\begin{verbatim}
  - Assignment A_\theta:
    x = true
    y = true
    z = false
    w = false

  - Refinement constraint \delta:
    x + y >= 1
    ^   ^
    |   |
    -----
      |
      S = {x, y}, therefore \delta excludes assignment A in the next iteration

@@

\end{verbatim}
\begin{align*}
  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
    \right) \right) \land \\
  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
\end{align*}
\begin{verbatim}

@@

\end{verbatim}
\begin{align*}
  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
\end{align*}
